---
title: "final"
format: pdf
---


```{r}
#| label: packages
library(tidyverse)
```

```{r}
#| label: data-sets
# creating new dataset
nests <- read.csv("nest_data_final.csv")
```

## Problem 1

### a.

In part 1 the test they used a logistic regression test.

In part 2 the test used was a Chi square test of independence.

### b. 

Part 1:
American Avocet microhabitat use was not determined by the distance of the microhabitat from the water.

Part 2:
The American Avocet, Forster's Tern, and Black-necked Stilts all occupy different habitats within the managed pond, non-tidal marsh, salt pond, tidal marsh, or tidal mudflat study areas.

### c.

One other variable that could influence the birds' habitat use is how deep the water is in different microhabitats.
This is a continuous quantitative variable that would be measured in meters. It could be measured with a meter stick at the different sites.
This could effect the American Avocet habitat use because the depth of the water could determine what prey live in the habitat and the resources available to the birds.














