---
title: "final"
format: pdf
---

## Set Up

```{r}
#| label: packages
library(tidyverse)
```

```{r}
#| label: data-sets
# creating new dataset
nests <- read.csv("nest_data_final.csv")
```

## Problem 1

### a.

In part 1 the test they used a logistic regression test.

In part 2 the test used was a Chi square test of independence.

### b. 

Part 1:
American Avocet microhabitat use was not determined by the distance of the microhabitat from the water.

Part 2:
The American Avocet, Forster's Tern, and Black-necked Stilts all occupy different habitats within the managed pond, non-tidal marsh, salt pond, tidal marsh, or tidal mudflat study areas.

### c.

One other variable that could influence the birds' habitat use is how deep the water is in different microhabitats.
This is a continuous quantitative variable that would be measured in meters. It could be measured with a meter stick at the different sites.
This could effect the American Avocet habitat use because the depth of the water could determine what prey live in the habitat and the resources available to the birds.


## Problem 2

### a. 

```{r}
#| label: problem-2a
# creating new object
nests_clean <- nests |> 
  # filtering to only include AB, RRB, and BBR
  filter(ant_species %in% c("AB", "RRB", "BBR")) |> 
  # Only using the 3 species
  mutate(
    ant_species = factor(
      ant_species,
      levels = c("AB",
                 "RRB",
                 "BBR")
    ),
    # creating new column of scientific names
    scientific_name = case_when(
      # adding in the scientific names for the species
      ant_species == "AB" ~ "C. sjostedti",
      ant_species == "RRB" ~ "C. mimosae",
      ant_species == "BBR" ~ "C. nigriceps"
    )) |> 
  # selecting columns of interest
  select(case_control, height_cm,
         scientific_name)
# showing structure of dataframe
str(nests_clean)
# displaying 10 rows
slice_sample(nests_clean, n = 10)
```


### b.

The 1s show when a tree had a nest in it and the 0 shows where there was an available tree with no nest.

### c.

Tree height is a continuous quantitative variable measured in cm. It could affect the probability of nest occurrence because as the tree height increases so does the probability of there being a nest also increases because it becomes more and more removed from predators. I found this in the papers introduction.

Ant species is a categorical variable of either C. sjostedi, C. mimosae, or C. nigriceps. The species of ants might effect the probability of a nest because they might have different ways of defending themselves from the birds that drives away the birds from making a nest. I found this in the end of the introduction.

### d.
### e.
### f.
### g.







